# Reproducible Research: Peer Assessment 1

## Setup required packages
```{r, echo=TRUE}
if ("ggplot2" %in% rownames(installed.packages()) == FALSE){
    install.packages("ggplot2")
}
library(ggplot2)
```

## Loading and preprocessing the data
```{r, echo=TRUE}
library(ggplot2)
unzip("activity.zip")
activityData <- read.csv("activity.csv", stringsAsFactors=F)
activityData$date <- as.Date(activityData$date)
cleanActivityData <- activityData[!is.na(activityData$steps),]
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
sums <- aggregate(cleanActivityData$steps, list(cleanActivityData$date), sum)
names(sums) <- c("date", "steps")
ggplot(sums, aes(date, steps)) + geom_histogram(stat="identity") +
    ggtitle("Steps taken per day") +
    xlab("Day") + ylab("Steps")

#Average total steps per day - original data
mean(sums$steps)

#Median total steps per day - original data
median(sums$steps)
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
averages <- aggregate(cleanActivityData$steps, list(cleanActivityData$interval), mean)
names(averages) <- c("interval", "interval.average")

ggplot(averages, aes(x=interval, y=interval.average)) + geom_line() +
    xlab("Interval") + ylab("Average steps walked") +
    ggtitle("Average steps walked per interval")

#Interval with highest average steps
averages[averages$interval.average == max(averages$interval.average), "interval"]
```

## Imputing missing values
### Imputation strategy - Replace NA values with average steps taken during interval across all days.
```{r, echo=TRUE}
#Merges average steps per interval (computed earlier) into source data 
activityData <- merge(activityData, averages, by="interval")

#Changes NA values to average steps per interval
activityData$steps[is.na(activityData$steps)] <-
    activityData$interval.average[is.na(activityData$steps)]

sums <- aggregate(activityData$steps, list(activityData$date), sum)
names(sums) <- c("date", "steps")

#Average total steps per day - imputed data
mean(sums$steps)

#Median total steps per day - imputed data
median(sums$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
activityData$day <- weekdays(activityData$date)
activityData$day[activityData$day %in% c("Saturday", "Sunday")] <- "Weekend"
activityData$day[activityData$day != "Weekend"] <- "Weekday"
activityData$day <- factor(activityData$day)

ggplot(activityData, aes(x=interval, y=steps)) + geom_line() +
    facet_grid(day ~ .) + xlab("Interval") + ylab("Steps") +
    ggtitle("Walking patterns by time of week")
```